Fishery Markets and Large-Scale Marine Conservation

### 2019-05-30 docs/old_ms/Markets_Conservation_May17.tex
Added all files

\begin{sciabstract}
Marine Protected Areas have the primary objective of conserving bounded waters by displacing or limiting fishing effort. With global conservation targets seeking to protect 10\% of the world’s ocean by 2020, it is important that the spatial displacement of economic activities is fully considered. A spatial closure in a trans-boundary fishery can potentially result in costs to the implementing country and benefits to neighboring countries. We show that effort markets may reduce up to 70\% of the costs of Marine Protected Areas, and that allocation rules based on biomass (rather than effort) may further incentivize conservation. We then use identification of fishing activity via Automatic Identification Systems to evaluate the effect that the Phoenix Islands Protected Area (PIPA, 397,447 km\textsuperscript{2}, implemented in 2015 in Kiribati) had on vessel distribution and behavior. Aggregate fishing effort remains relatively constant following the closure, but is significantly reduced within Kiribati, especially by vessels that historically fished within PIPA before its implementation. We find evidence that Kiribati may be trading its surplus of vessel-days, which would significantly reduce the costs of displaced fishing effort. We use these insights to predict the implications of a proposed Large-Scale Marine Protected Area (LSMPA) in Palau (another PNA member) and estimate that effort markets could reduce the cost of implementation by as much as 77\%. Our work highlights the role of effort markets in enabling conservation, and the importance of allocation rules that praise, not punish, conservation. The Parties to the Nauru Agreement have successfully developed a management scheme that allows for conservation interventions to come at a low cost, and could be a model for trans-boundary conservation across the globe.
\end{sciabstract}

@@

\section{Introduction}\label{introduction}

>>>>

\begin{sciabstract}
Will rights-based approaches to managing natural resources facilitate, or relegate large-scale conservation? We study this question in the context of large-scale marine protected areas, and ask whether a country who commits to large-scale conservation is destined to suffer commensurate losses in fishery income.  Indeed, this is the question island (and other) nations all over the world face as they consider whether, and how, to comply with various international commitments to massively expand marine protection. We use a combination of spatial bioeconomic modeling, which accounts for fisheries markets such as the vessel day trading scheme in the Pacific, and vessel-level empirical data on fishing behavior before and after MPA implementation, to assess how market design incentivizes, or relegates conservation. The answer hinges critically on two design features of the rights-based fishery market. First, only if trading of harvest or vessel days is allowed across countries can a conservation-minded country capture the economic spillover benefits of their conservation.  Instead, if trading is forbidden, then countries are likely to suffer significant losses from conservation. Second, as rights are re-allocated between countries periodically, the re-allocation must depend only on biological features, and not on a country's past effort or catch. Otherwise, the reallocation acts as a punishment to the conserving country. Overall, these results show that seemingly innocuous design features of a market can have indelible impacts on a country's willingness to engage in large-scale conservation; these results are confirmed with an empirical case study from the Phoenix Islands Protected Area, one of the world's largest marine reserves.
\end{sciabstract}

@@

\section{Introduction}\label{introduction}
Recognizing a need to secure biodiversity, blue carbon, rare habitats, and other ecosystem services, various international bodies have committed to dramatically expand marine protection around the world. Currently only about 3\% of the world's ocean is strongly protected, and these commitments call for 10\%, 30\%, or even up to 50\% of the ocean to be off-limits to fishing. Fundamentally, to achieve these goals, huge swaths of sovereign nations' waters will have to be closed to fishing. What incentives might motivate a country to engage in such large-scale marine protection?  Would any country rationally close 20\%, 50\%, or even 100\% of its waters to fishing if this meant losing all fishing revenue from its borders?  We show that a country's incentives for large-scale conservation hinge critically on an unsuspected culprit: The design of fishing rights.

We are motivated by a real-world institution called the Parties to the Nauru Agreement (PNA), which is one of the most successful, profitable, and well-orchestrated international cartels from any sector of the economy. Like an OPEC for tuna, the PNA is a nine-country coalition of Pacific Island states that collectively manages tuna fishing in its members' waters.  These waters collectively account for XXX square km (an area larger than XXX), and YYY\% of the catch of productive tuna species such as XXX and YYY.

In addition to highly productive tuna fisheries, the PNA waters provide a wealth of other ecosystem benefits; this explains why international attention has focused there for targeted conservation efforts to help achieve the previously mentioned targets. But the PNA member nations derive enormous benefits from leasing tuna fishing rights to foreign fleets, often exceeding 50\% of GDP. A natural concern arises: Even if the closure creates net benefits to the world, and perhaps even to tuna fisheries Pacific-wide, the economic loss to the conserving country could still be enormous, and may prevent it from engaging in any meaningful conservation. Intuitively, then, it seems that a country's willingness to engage in large-scale conservation will depend crucially on whether it can successfully capture a sufficiently high fraction of the benefits that are derived from that conservation.  We show how the rights-based fisheries market can be designed (if it is new) or modified (if it already exists) to ensure that the conserving country is rewarded, not punished, for large-scale conservation.

OLD TEXT FROM HERE
